The Galactic Midplane Is Not a Plane: Implications for Dynamical Analysis with \textit{Gaia} Data and Beyond

### 2019-05-08 paper/ms.tex
Update ms.tex
Fix travis
Small updates

\author{Chervin F. P. Laporte}
\altaffiliation{Simons Fellow}
\affil{\victoria}

@@

\subsection{Description of Milky Way-Sagittarius Interaction Simulation} 
\label{ssec:sag_sim}
In addition to the cosmological zoom-ins, we will also briefly consider results
from a toy Sagittarius encounter simulation. This simulation offers us the
ability to see how the midplane varies in a more controlled environment. The
details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
briefly summarize the most relevant details here. For the Milky Way, the dark
halo is modeled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
length of $52\,\kpc$, the disk modeled as an exponential disk with a scale
radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
$6\times10^{10}\,\Msun$, a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. Sagittarius is modeled as a
dark matter Hernquist sphere of mass \Gus{something} and scale length
\Gus{something}, along with a stellar component modeled as a Hernquist sphere
of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
are realized with live distributions of N-body particles; the Milky Way and Sagittarius are
each initialized to be in equilibrium in isolation \Gus{check!}.

@@

\subsection{The Local Midplane} \label{ssec:local_midplane}
Using the two sets of simulations, we determine the local midplane that an
observer might measure if they were situated in each of these galaxies as a
function of azimuth at the solar circle. Starting from the coordinate system
described in the previous section, which is approximately aligned so that the
$z$-coordinate is perpendicular to the disk plane, we place our imaginary
observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
between $0<\phi<2\pi$.\footnote{The immortal observer also has warp drive, allowing them to travel around the galaxy much more quickly than its dynamical timescale (warp factor $\sim0.25$). Using warp drive allows them to sidestep time dilation effects.} At
each value of $\phi$ we then compute the median $z$ for stars within a
cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
fiducial disk. We then re-define the new midplane of the cylinder to be the
median $z$, re-select stars, and iterate until the median $z$ value converges.
We find that only $10$ iterations of this procedure are necessary for
convergence. The resulting median $z$ is taken to be what our observer would
measure as the local Galactic midplane at each $\phi$.

@@

\acknowledgments
We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
would like to thank Todd Phillips for helpful discussions. A.B. was supported
in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
the Roy \& Diana Vagelos Challenge Award. K.V.J.'s work was performed in part
during the Gaia19 workshop and the 2019 Santa Barbara Gaia Sprint (also
supported by the Heising-Simons Foundation), both hosted by the Kavli
Institute for Theoretical Physics at the University of California, Santa
Barbara. K.V.J.'s contributions were supported in part by the National Science
Foundation under grants NSF PHY-1748958 and NSF AST-1614743. M.-M.M.L. was
partly supported by NSF grant AST18-15461. The work of D.A.-A., A.B., D.W.H.,
C.F.P.L., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
Foundation.

>>>>

\author{Chervin F. P. Laporte}
\altaffiliation{CITA National Fellow}
\affil{\victoria}

@@

\subsection{Description of Milky Way-Sagittarius Interaction Simulation} 
\label{ssec:sag_sim}
In addition to the cosmological zoom-ins, we will also briefly consider results
from a live N-body simulation of a Sagittarius-like encounter. This simulation offers us the
ability to see how the midplane varies in a more controlled environment. The
details of the simulation are given in \citet{2018MNRAS.481..286L}, but we
briefly summarize the most relevant details here. For the Milky Way, the dark
halo is modeled as a Hernquist sphere of mass $10^{12}\,\Msun$ and scale
length of $52\,\kpc$, the disk modeled as an exponential disk with a scale
radius of $3.5\,\kpc$, scale height $0.53\,\kpc$, and mass
$6\times10^{10}\,\Msun$, and a Hernquist bulge of mass $10^{10}\,\Msun$ and scale
length $0.7\,\kpc$ \citep{1990ApJ...356..359H}. We use the L2 model, for which Sagittarius is modeled as a
dark matter Hernquist sphere of mass $8\times10^{10}\,\Msun$ and scale length
$8\,\kpc$, along with a stellar component modeled as a Hernquist sphere
of mass $6.4\times10^8\,\Msun$ and scale radius $0.85\,\kpc$. All components
are realized with live distributions of N-body particles; the Milky Way and Sagittarius are
each initialized to be in equilibrium in isolation \Gus{check!}.

@@

\subsection{The Local Midplane} \label{ssec:local_midplane}
Using the two sets of simulations, we determine the local midplane that an
observer might measure if they were situated in each of these galaxies as a
function of azimuth at the solar circle. Starting from the coordinate system
described in the previous section, which is approximately aligned so that the
$z$-coordinate is perpendicular to the disk plane, we place our imaginary
observer at $z=0$ and a galactocentric cylindrical radius of $8.2\,\kpc$ and vary the azimuth
between $0<\phi<2\pi$. At
each value of $\phi$ we then compute the median $z$ for stars within a
cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$ perpendicular to the
fiducial disk. We then re-define the new midplane of the cylinder to be the
median $z$, re-select stars, and iterate until the median $z$ value converges.
We find that only $10$ iterations of this procedure are necessary for
convergence. The resulting median $z$ is taken to be what our observer would
measure as the local Galactic midplane at each $\phi$.

@@

\acknowledgments
We would like to thank Megan Bedell, Robert A. Benjamin, Tobias Buck, Elena
D'Onghia, Benoit Famaey, and Adrian Price-Whelan for helpful discussions. A.B.
would like to thank Todd Phillips for helpful discussions. A.B. was supported
in part by the Roy \& Diana Vagelos Program in the Molecular Life Sciences and
the Roy \& Diana Vagelos Challenge Award. K.V.J.'s and C.F.P.L.'s work was
performed in part during the Gaia19 workshop and the 2019 Santa Barbara Gaia
Sprint (also supported by the Heising-Simons Foundation), both hosted by the
Kavli Institute for Theoretical Physics at the University of California, Santa
Barbara. This work used the Extreme Science and Engineering Discovery
Environment (XSEDE), which is supported by National Science Foundation grant
number OCI-1053575. K.V.J.'s contributions were supported in part by the
National Science Foundation under grants NSF PHY-1748958 and NSF AST-1614743.
M.-M.M.L. was partly supported by NSF grant AST18-15461. The work of D.A.-A.,
A.B., D.W.H., R.S., K.V.J., M.-M.M.L., and M.K.N. is supported by the Simons
Foundation.
